The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-05 writing/draft3/draft7.tex
writing

In this study, we analyzed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
	We acquired the county level yield data from the Famine Early Warning Systems Network (FEWS NET). As for the weather data, we utilized $0.25^\circ$ resolution precipitation and temperature gridded datasets. The precipitation data come from the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth.

	Both precipitation and temperature datasets are daily while the yield data are yearly by its nature. Hence, in order to proceed with the analysis, the daily precipitation and temperature data had to be converted
into a county-level yearly panel. The first step of this procedure was averaging the gridded weather data over the counties of Kenya resulting in a single daily time series for each county. In the next step, we aggregated the resulting county level time series temporally to obtain yearly time series using the seasonal crop calendar which is available at the website of FEWS~NET. We separated the counties of Kenya into two groups, such that the crop calendar can be considered homogeneous within each group. The groups are: \textit{(i)} Arid and semi-arid (ASAL) counties and \textit{(ii)} non-ASAL counties as defined by the ASAL Stakeholders Forum (ASF). For the list of counties in each of the groups see table~\ref{ASAL} in the appendix. \footnote{The crop calendar is probably not exactly the same for all counties within each group, however, it is similar enough so that it can be considered homogeneous within the groups. To a certain extent, the crop calendar is likely to differ even within a single county. However, considering the information provided by ASF, FEWS NET and accounting for our intermediate results, we opted for the two groups as a compromise between accuracy and complexity.}.

In terms of weather, the most relevant periods for yield are the rainy seasons.
Therefore, we only considered weather during the rainy seasons for our analysis. According to FEWS~NET, the ASAL counties have two distinct rainy seasons while the non-ASAL counties have one longer rainy seasons. The two rainy seasons of the ASAL counties are: \textit{(i)} March-May and \textit{(ii)} October-December. In the non-ASAL area, the rainy season is in March-August.

Hence, for the ASAL counties, we aggregated daily precipitation and temperature over October, November and December of the previous year and May, April, May of the current year  to get single value of the weather variables for each county and year. For the non-ASAL counties we aggregated the daily weather values over March, April, May, June, July and August to obtain yearly values.

\subsection{Statistical approach}

% No need to write too many details (or a section) about measures of food security
% No need to describe all the weather measures/characteristics that we have calculated and/or tried to use and didn't work
\begin{itemize}
\color{blue}
\item \textit{We used commonly used measures of weather/drought (Only mention the significant weather measures/variables)}
\item Describe the temporal aggregation of the weather variables, seasons
\end{itemize}
\color{black}

>>>>

In this study, we analyzed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
	We acquired the county level yearly yield data from the Famine Early Warning Systems Network (FEWS NET). As for the weather data, we utilized $0.25^\circ$ resolution precipitation and temperature gridded datasets. The precipitation data were obtained from the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth. We calculated spatial averages of the gridded data over the counties resulting in a single value for each county and each point in time. The frequency of precipitation and temperature data was daily, therefore, further aggregation was needed to obtain yearly values corresponding to the yearly frequency of yield data. Hence, each weather characteristic was further aggregated over years resulting in a county-level panel conformable with the yield data.

	There are two predominant rainfall regimes in Kenya. The Arid and semi-arid (ASAL) counties mostly exhibit bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October to December. In the non-ASAL counties, the single rainy season usually starts in March and lasts until August. Following the precipitation regimes and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.

Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature. Only some of them were found significant. These include:
\subsection{Statistical approach}

% No need to write too many details (or a section) about measures of food security
% No need to describe all the weather measures/characteristics that we have calculated and/or tried to use and didn't work
